Title

### 2022-03-18 paper/MDPI_template/template.tex
Updated paper

%=================================================================
\documentclass[sensors,article,submit,pdftex,moreauthors]{Definitions/mdpi} 
% For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept". The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}. This is especially recommended for submission to arXiv, where line numbers should be removed before posting. For preprints.org, the editorial staff will make this change immediately prior to posting.

@@

%=================================================================
% Full title of the paper (Capitalized)
\Title{A Deep Learning Approach for Gait Event Detection from a Single Shank-Worn IMU: Validation in Healthy and Neurological Cohorts}

% MDPI internal command: Title for citation in the left column
\TitleCitation{A Deep Learning Approach for Gait Event Detection from a Single Shank-Worn IMU: Validation in Healthy and Neurological Cohorts}

% Author Orchid ID: enter ID or remove command
\newcommand{\orcidauthorA}{0000-0002-2507-0924} % Add \orcidA{} behind the author's name
%\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name

% Authors, for the paper (add full first names)
\Author{Robbin Romijnders$^{1}$\orcidA{}, Elke Warmerdam$^{2}$, Clint Hansen$^{1}$*, Gerhard Schmidt$^{3}$ and Walter Maetzler$^{1,}$}

@@

% MDPI internal command: Authors, for metadata in PDF
\AuthorNames{Robbin Romijnders, Elke Warmerdam, Clint Hansen, Gerhard Schmidt and Walter Maetzler}

% MDPI internal command: Authors, for citation in the left column
\AuthorCitation{Romijnders, R.; Warmerdam, E.; Hansen, C.; Schmidt, G.; Maetzler, W.}
% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Department of Neurology, Kiel University, 24105 Kiel, Germany; r.romijnders@neurologie.uni-kiel.de (R.R.), c.hansen@neurologie.uni-kiel.de (C.H.), w.maetzler@neurologie.uni-kiel.de (W.M.)\\
$^{2}$ \quad xxxxx, xxxxx, xxxxx Homburg, Germany; elke.warmerdam@uni-saarland.de (E.W.)\\
$^{3}$ \quad Institute of Electrical Engineering and Information Technology, Faculty of Engineering, Kiel University, 24143 Kiel, Germany; gus@tf.uni-kiel.de (G.S.)}

% Contact information of the corresponding author
\corres{Correspondence: r.romijnders@neurologie.uni-kiel.de (R.R.)}

% Current address and/or shared authorship
%\firstnote{Current address: Affiliation 3.} 
%\secondnote{These authors contributed equally to this work.}
% The commands \thirdnote{} till \eighthnote{} are available for further notes

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Gait deficits are common in older adults and possibly reflect the presence of an underlying neurodegenerative disease \cite{Snijders2007,Hodgins2008}. For example, conversion to Parkinson's Disease \cite{DelDin2019} or from mild cognitive impairment to Alzheimer's Disease \cite{Koenig2017,Bertoli2018} are linked with changes in spatiotemporal gait parameters. Similarly, temporal gait parameters are different for stroke patients \cite{SchroederVon1995,Mohan2021} and patients with multiple sclerosis \cite{Griskevicius2016,Flachenecker2019} when compared to healthy controls. To objectively quantify gait deficits, stride-specific parameters such as stride time or stride length are often used \cite{Hannink2016}, where the beginning and end of a stride are determined from two successive initial contacts (ICs) of the same foot \cite{Perry2010,Whittle2012}. The IC is when the foot contacts the ground and together with the instant at which the foot leaves the ground (final contact, FC), each stride can be divided in a stance and swing phase \cite{Rueterbories2010,Bruening2014}.  The events of IC and FC, also referred to as \emph{gait events}, are commonly determined using force or pressure measuring devices \cite{Bruening2014}, or stereophotogrammetry systems based on reflective markers using a multi-camera setup \cite{Chiari2005}. These systems are relatively expensive, and restricted to usage in expertise laboratories \cite{Iosa2016,Jarchi2018}. As there is increasing evidence that gait measured in the lab does not reflect daily-life gait \cite{Hillel2019,Warmerdam2020,Atrsaei2021}, there is more and more interest in measurement systems that allow for continuous gait analysis in ambulatory settings. Therefore, the use of inertial measurement units (IMUs) is especially attractive, as these can be used to measure gait in ecologically valid environments, such as the home environment, thereby painting a more complete picture of health status \cite{DelDin2016,Shah2020} and providing clinical information that is complementary to standardized lab-based assessments \cite{Warmerdam2020,Fasano2020,Atrsaei2021,Corra2021}.

Previous research suggests that gait event detection is more accurate using an IMU worn on a lower limb (e.g., shank or foot) compared to an IMU worn on the low back \cite{BenMansour2015,Storm2016,Panebianco2018}. Now, in order to get from abstract IMU sensor readings to clinically relevant gait parameters (e.g., from accelerations and angular velocities to stride times) \cite{Hannink2016}, different algorithmic approaches have been developed in the last twenty years of clinical gait research. A recent study has evaluated a cross-section of these algorithms for different sensor locations on the lower leg and foot \cite{Niswander2021}. The algorithms were categorized according to which signals were analyzed, for example the angular velocity about the medio-lateral axis, or the accelerations along vertical and antero-posterior axis. This means the sensor readings need to be linked with the anatomical axes, that is, one needs to know which sensor axis aligns with for example the medio-lateral axis. In most approaches, it is simply assumed that due to sensor attachment the sensor axis aligns roughly with the anatomical axis of interest (\cite{Salarian2004,Sabatini2005,Jasiewicz2006,Catalfamo2010,Trojaniello2014,Maqbool2016,Romijnders2021}) or an additional calibration procedure (e.g., \cite{Ferraris1995}) is required (\cite{Greene2010,Niswander2021}). In ambulatory assessments however, study participants often attach the sensor themselves, and therefore the sensor location and alignment cannot be controlled for. Furthermore, it is unlikely that each time the sensor is (re-)attached study participants, especially those with gait deficits, perform a calibration procedure that usually consists of holding a pre-defined pose and performing some known movement sequences (TODO: reference?! Kong? Seel?).

Taken together, this drives the need for an approach that is invariant to sensor orientation, and is applicable across a variety of pathological gait patterns. As pointed out by \cite{Hannink2016}, in the field of image analysis similar requirements have been successfully addressed by algorithms that share a common underlying methodology referred to as deep learning \cite{LeCun2015,TerHaarRomeny2019}. Recent applications of deep learning algorithms have already shown improved performance in detecting gait events from marker-based motion capture when compared to conventional, often heuristics-based, algorithms \cite{Kidzinski2019,Lempereur2020,Filtjens2020}. Another study used a deep learning approach to detect gait events from either three IMUs (worn on the low back, and both ankles) or a single IMU (worn on the low back), and showed that the time error was considerably smaller for the deep learning algorithm than for a commonly applied wavelet-based approach \cite{Gadaleta2019}.

In this study, we further these works by validating this deep learning approach for detecting gait events in a heterogeneous cohort of healthy and neurologically diseased adults considering a single IMU setup, worn on the lower leg.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Materials and Methods}

\subsection{Data Collection}
Gait analyses were performed in the Universit\"{a}tsklinikum Schleswig-Holstein (UKSH) campus Kiel, Germany. The study \cite{Warmerdam2021} was approved by the ethical committee of the medical faculty at the UKSH (no: D438/18). In total, data from 160 participants were included for the current analysis, including data from young adults (YA; age: 18 - 60 years), older adults (OA; age: $>$60 years), people with Parkinson's Disease (PD; according to the UK Brain Bank criteria \cite{Gibb1988}), people with a recent ($<$4 weeks) symptomatic stroke (stroke), people with multiple sclerosis (MS; according to the McDonalds criteria \cite{Thompson2018}), people with chronic low back pain (cLBP), and people with other diagnoses that were assumed not to affect mobility (Table \ref{tab:demographics_data}). Inclusion criteria were an age of 18 years or older, and the ability to walk independently without a walking aid. Participants were excluded from the study with a Montreal Cognitive Assessment \cite{Nasreddine2005} score $<$15 and other movement disorders that affected mobility, as noticed by the clinical assessor.

\begin{table}[H] 
	\caption{\label{tab:demographics_data}Demographics data of the study participants. Age, height, and weight are presented as mean (standard deviation).}
	\newcolumntype{C}{>{\centering\arraybackslash}X}
	\begin{tabularx}{\textwidth}{CCCCCC}
		\toprule
		\textbf{Group}	& \textbf{Gender}	& \textbf{Number of}	& \textbf{Age}	& \textbf{Height}	& \textbf{Weight}\\
		 & & \textbf{subjects} & years	& cm	& kg\\
		\midrule
		YA		& F		& 21	& 27 (7)	& 173 (5)	& 67 (8)\\
		 & M	& 22	& 30 (9)	& 185 (8) & 81 (13)\\ 
		\midrule
		OA		& F		& 11	& 70 (6)	& 167 (6)	& 72 (17)\\
		 & M	& 11	& 73 (6)	& 180 (6) & 83 (11)\\
		\midrule
		PD		& F		& 12	& 67 (6)	& 168 (6)	& 70 (14)\\
		 & M	& 20	& 62 (11)	& 178 (6) & 87 (13)\\
		\midrule
		MS		& F		& 12	& 37 (9)	& 174 (9)	& 75 (9)\\
		 & M	& 9	& 42 (15)	& 189 (8) & 96 (30)\\
		\midrule
		stroke	& F		& 4	& 66 (10)	& 160 (6)	& 65 (11)\\
		 & M	& 17	& 67 (17)	& 178 (7) & 84 (14)\\
		\midrule
		cLBP	& F		& 3	& 64 (10)	& 166 (5)	& 65 (5)\\
		 & M	& 7	& 63 (16)	& 178 (7) & 90 (16)\\ 
		\midrule
		other	& F		& 3	& 60 (13)	& 166 (3)	& 79 (15)\\
		 & M	& 8	& 68 (18)	& 182 (7) & 85 (13)\\ 
		\bottomrule
	\end{tabularx}
	\noindent{\footnotesize{\textbf{Group}: YA: younger adults, OA: older adults, PD: Parkinson's Disease, MS: multiple sclerosis, cLBP: chronic low back pain; \textbf{Gender}: F: female, M: male.}}
\end{table}

Participants performed three walking trials consisting of walking 5 meter at either (1) preferred speed ("\emph{Please walk at your normal walking speed.}"), (2) slow speed ("\emph{Please walk half of your normal walking speed.}"), or (3) fast speed ("\emph{Please walk as fast as possible, without running or falling.}"). The 5 meter distance was marked with two cones on both ends, and participants were asked to start walking approximately two steps before the cones on one end, and stop walking approximately two steps after passing the cones on the other end.

For the current analysis data from four IMUs (Noraxon USA Inc., myoMOTION, Scottsdale, AZ, USA) were considered, namely those that were attached laterally above the left and right ankle joint and those attached proximally at the left and right shank. IMUs were secured to participants using elastic bands with a special hold for the IMU. Furthermore, reflective markers were attached on top of the usual foot wear at the heel and toe of both feet. Marker data were recorded using a twelve-camera stereophotogrammetry system (Qualisys AB, G\"{o}teborg, Sweden) at a sampling frequency of 200 Hz. IMU data were recorded at the same sampling frequency, and both systems were synchronized using TTL signal \cite{Warmerdam2021}. For some recordings, the sampling frequency of the IMUs was erroneously set at 100 Hz, and therefore data from these trials were upsampled before further analysis.

\subsection{Data Pre-processing \label{subsec:data_preprocessing}}
\subsubsection{Reference Values}
The raw reflective marker data were first used to determine the start and end of each trial. The start of the trial was defined as the time for which the first toe marker (left or right) crossed the \emph{virtual} starting line. The end was defined as the time for which the last heel marker crossed the end line. For both marker and IMU systems, data were then cropped from start to the end of the trial.

Any remaining gaps in the marker data were filled by interpolation making use of inter-correlations between markers \cite{Federolf2013,Gloersen2016}. The data were then low-pass filtered using a double pass 6th order Butterworth filter with a cut-off frequency of 20 Hz \cite{Kormylo1974,Racz2021}. The filtered data were differentiated to get velocity signals, and timings of ICs and FCs were determined from local maxima and minima in the heel and toe vertical velocity signals \cite{Pijnappels2001,OConnor2007}. Like in \cite{Carcreff2018,Romijnders2021}, all identified ICs and FCs were manually checked using Qualisys Track Manager 2018.1 software (QTM; Qualisys AB, Göteborg, Sweden).

\subsubsection{IMU Data}
For the deep learning model, the data from randomly selected one-third of participants were used as a "training set", another one-third was used as a "validation set", and the rest was used as "test set", where randomization was stratified by both group and gender. Accelerometer and gyroscope data from the four IMUs were normalized by subtracting the channel-wise mean, and dividing by the channel-wise standard deviation. Then, for the training and validation set, the time series data were partitioned into equal length time windows of 400 samples, with an overlap of 50\% between successive windows.

\subsection{Model}
\subsubsection{Notation}
Data from the IMUs' 3D accelerometer and 3D gyroscope were used. A signal was considered a sequence of real-valued numbers: $\mathbf{x} = \begin{array}{ccccc}
	\cdots & x[n-1] & x[n] & x[n+1] & \cdots
\end{array}$, with $n = 1, \cdots, N$ the discrete-time sample index. For any given channel, e.g., acceleration in $X$-direction, the signal was denoted in vector notation as: $\mathbf{x}_{d} = \begin{bmatrix}
x_{d}[1] & x_{d}[2] & \cdots & x_{d}[N]
\end{bmatrix}^{\mathrm{T}}$, where $d$ referred to the $d$-th channel.

Data from all channels of a single IMU were collected in a $ N \times D$ data matrix, like:
\begin{equation}
	\mathbf{X} = \begin{bmatrix}
		\vdots & \vdots &  & \vdots\\
		\mathbf{x}_{1} & \mathbf{x}_{2} & \cdots & \mathbf{x}_{D}\\
		\vdots & \vdots &  & \vdots
	\end{bmatrix} = \begin{bmatrix}
		x_{1}[1] & x_{2}[1] & & x_{D}[1]\\
		x_{1}[2] & x_{2}[2] & & x_{D}[2]\\
		\vdots & \vdots & \cdots & \vdots\\
		x_{1}[N] & x_{2}[1] & & x_{D}[N]\end{bmatrix}, \quad \mathbf{X} \in \mathbb{R}^{N \times D}
	\label{eqn:train_data_single_instance}
\end{equation}

The corresponding labels (or targets) were then denoted as:
\begin{equation}
	\mathbf{y}_{\mathrm{IC}} = \begin{bmatrix}
		y_{1}[1]\\
		y_{1}[2]\\
		\vdots\\
		y_{1}[N]
	\end{bmatrix}, \quad \mathbf{y}_{\mathrm{FC}} = \begin{bmatrix}
		y_{2}[1]\\
		y_{2}[2]\\
		\vdots\\
		y_{2}[N]
	\end{bmatrix}, \quad y_{i}[n] \in [0, 1]
	\label{eqn:train_labels_single_instace}
\end{equation}

\subsubsection{Model Architecture}
The basic architecture for the deep learning model was a temporal convolutional network (TCN) \cite{Bai2018,Remy2020,Filtjens2020}. The TCN consisted of repeating blocks of dilated convolutions (CONV) \cite{YuKoltun2016,Bai2018} that were followed by batch normalization (BN) \cite{Ioffe2015}, rectified linear unit (ReLU) activation, and dropout (DropOut) \cite{Srivastava2014}. For each given dilation factor, the sequence of layers CONV-BN-ReLU-DropOut was repeated twice \cite{Remy2020}. Dilation factors were given as a sequence of increasing powers of 2, e.g., $\left\{1, 2, 4, 8\right\}$ \cite{VanDenOord2016,YuKoltun2016,Bai2018}. These repeating blocks were followed by a fully-connected (dense) layer with sigmoid activation, and outputs were predicted separately for ICs and FCs. The mean squared error (MSE) was used as loss function, and a gradient descent-based optimization algorithm with adaptive moment (Adam) optimizer was used to iteratively learn the weights \cite{Kingma2014,Schmidt2021}.

\subsubsection{Hyperparameter Optimization}
In order to find the best model architecture, hyperparameter tuning was perfomed using KerasTuner \cite{OMalley2019}. Here, the number of filters, the kernel size, and the maximum dilation factor (Table \ref{tab:hyperparameter_optimization}) were optimized for using a random search strategy \cite{Bergstra2012}.

\begin{table}[H] 
	\caption{Model hyperparameters that were optimized for, and the corresponding sets of possible values.\label{tab:hyperparameter_optimization}}
	\newcolumntype{C}{>{\centering\arraybackslash}X}
	\begin{tabularx}{\textwidth}{CC}
		\toprule
		\textbf{Description}	& \textbf{Possible values}\\
		\midrule
		Number of filters	& 8, 16, 32, 64, 128\\
		Kernel size			& 3, 5, 7\\
		Dilations 			& [1, 2], [1, 2, 4], [1, 2, 4, 8]\\
		\bottomrule
	\end{tabularx}
\end{table}

The model architecture that resulted from the hyperparameter optimization was then trained on the combined set of training and validation data. The trained model was used to predict occurrence of gait events from the hold-out test set.

\subsection{Analysis}
The predictions of the model on the test set data were compared with the labels from the test set. The model performance was evaluated for (1) overall detection performance, (2) time agreement between the predicted events and the (marker-based) annotated events, (3) agreement between subsequently derived gait parameters.
\subsubsection{Overall detection performance \label{subsubsec:overall_detection_performance}}
The overall detection performance quantified how many of the annotated evens were detected by the model (true positives, TP), how many of the annotated events were not detected (false negatives, FN), and how many event that were detected, were actually not annotated (false positives, FP). From these metrics, the recall (or sensitivity), precision and F$_{1}$ score were calculated as:
\begin{equation}
	\textrm{recall} = \frac{\textrm{TP}}{\textrm{TP} + \textrm{FN}} \nonumber
	\label{eqn:recall}
\end{equation}
\begin{equation}
	\textrm{precision} = \frac{\textrm{TP}}{\textrm{TP} + \textrm{FP}} \nonumber
	\label{eqn:precision}
\end{equation}
\begin{equation}
	\textrm{F$_{1}$ score} = 2 \cdot \frac{\textrm{recall} \cdot \textrm{precision}}{\textrm{recall} + \textrm{precision}}  \nonumber
	\label{eqn:f1_score}
\end{equation}

\subsubsection{Time agreement}
For all correctly detected gait events (TP, Section \ref{subsubsec:overall_detection_performance}), the time error between the annotated and detect gait event was defined as:
\begin{equation}
	\textrm{time error} = t_{\textrm{ref}} - t_{\textrm{pred}}
	\label{eqn:time_agreement}
\end{equation}
with $t_{\mathrm{ref}}$ the gait event time from the marker-based annotations, and $t_{\mathrm{pred}}$ the gait event time from the model predictions. As a robust measure for the average time error and its spread, the median time error and the inter-quartile range (IQR) were reported \cite{OpenIntro2019}.

\subsubsection{Gait parameters}
For those trials for which all gait events were detected, and no spurious events were detected, the following stride parameters were calculated:
\begin{equation}
	\textrm{stride time} = t_{\textrm{IC}_{k}} - t_{\textrm{IC}_{k-1}}
	\label{eqn:stride_time}
\end{equation}
\begin{equation}
	\textrm{stance time} = t_{\textrm{FC}_{k}} - t_{\textrm{IC}_{k-1}}
	\label{eqn:stance_time}
\end{equation}
\begin{equation}
	\textrm{swing time} = t_{\textrm{IC}_{k}} - t_{\textrm{FC}_{k}}
	\label{eqn:swing_time}
\end{equation}

@@

\subsection{Overall detection performance}
The performance of detecting initial contacts and final contacts was objectively quantified by the number of annotated events that were detected (true positives, TP), the number of annotated events that were not detected (false negatives, FN), and the number of detected events that were not annotated (false positives, FP). From these, recall, precision and F$_{1}$ score were calculated (Table \ref{tab:overall_detection_performance}).
\begin{table}[H]
	\caption{Overall detection performance for initial contacts and final contacts as quantified by recall, precision and F$_{1}$ score.\label{tab:overall_detection_performance}}
	\begin{adjustwidth}{-\extralength}{0cm}
		\newcolumntype{C}{>{\centering\arraybackslash}X}
		\begin{tabularx}{\fulllength}{r|CCCCCC|CCCCCC}
			\toprule
			 & \multicolumn{6}{c}{\textbf{Initial contacts}}	& \multicolumn{6}{c}{\textbf{Final contacts}}\\
			\textbf{Tracked point}	& \textbf{TP}	& \textbf{FN}	& \textbf{FP}	& \textbf{recall}	& \textbf{precision}	& \textbf{F1}	& \textbf{TP}	& \textbf{FN}	& \textbf{FP}	& \textbf{recall}	& \textbf{precision} 	& \textbf{F1}\\
			\midrule
			left ankle		& 624	& 19	& 5		& 97\%	& 99\%	& 98\%		& 606	& 32	& 10	& 95\%	& 98\%	& 97\%\\
			right ankle		& 599	& 42	& 8 	& 93\%	& 99\%	& 96\%		& 614	& 17	& 12	& 97\%	& 98\%	& 98\%\\
			left shank		& 605	& 38	& 15	& 94\%	& 98\%	& 96\%		& 585	& 53	& 18	& 92\%	& 97\%	& 94\%\\
			right shank		& 603	& 36	& 15 	& 94\%	& 98\%	& 96\%		& 595	& 30	& 9		& 95\%	& 99\%	& 97\%\\
			\bottomrule
		\end{tabularx}
	\end{adjustwidth}
	\noindent{\footnotesize{TP: true positives, FN: false negatives, FP: false positives, F1: F$_{1}$ score.}}
\end{table}

For both ICs and FCs, recall is high for each of the tracked points (i.e, $\ge 92\%$), and so is precision (i.e, $\ge 97\%$). Differences between the tracked points are small, i.e. the minimum recall is $92\%$ and the maximum recall is $97\%$, and the minimum precision is $97\%$ and the maximum precision is $99\%$.

\subsection{Time error}
\begin{figure}[H]
	\includegraphics[width=10.5 cm]{fig/box_plots_tracked_points_with_outliers.png}
	\caption{Time errors for initial (left) and final (right) contacts detection, for each of the different tracked points.\label{fig:time_error_box_plots}}
\end{figure}

Time errors for ICs and FCs are visually depicted using boxplots for each tracked point. Data were not normally distributed, thus Wilcoxon signed-rank tests were used to evaluate whether the median time error was zero.
\begin{table}[H] 
	\caption{Time errors for the correctly detected gait events. \label{tab:time_errors}}
	\newcolumntype{C}{>{\centering\arraybackslash}X}
	\begin{tabularx}{\textwidth}{r|CCC|CCC}
		\toprule
		 & \multicolumn{3}{c}{\textbf{Initial contacts}} & \multicolumn{3}{c}{\textbf{Final contacts}}\\
		\textbf{Tracked point} & \textbf{median} & \textbf{IQR} & (\textbf{Q25}, \textbf{Q75}) & \textbf{median} & \textbf{IQR} & (\textbf{Q25}, \textbf{Q75})\\
		 & ms & ms &  (ms, ms) & ms & ms & (ms, ms)\\
		\midrule
		left ankle & 0 & 20 & (-10, 10) & 0 & 10 & (-5, 5)\\
		right ankle & 0 & 20 & (-15, 5) & -5 & 15 & (-10, 5)\\
		left shank & -5 & 20 & (-10, 10) & -5 & 20 & (-15, 5)\\
		right shank & -2.5 & 20 & (-15, 5) & -5 & 20 & (-15, 5)\\
		\bottomrule
	\end{tabularx}
	\noindent{\footnotesize{IQR: inter-quartile range, Q25: 25th percentile, Q75: 75th percentile.}}
\end{table}

\subsection{Gait parameters}
\begin{figure}[H]
	\begin{adjustwidth}{-\extralength}{0cm}
		\centering
		\includegraphics[width=13.5cm]{fig/bland_altman_plots_stride_params.png}
	\end{adjustwidth}
	\caption{The agreement of extracted gait parameters between the sensor-based and marker-based methods.\label{fig:gait_parameters_bland_altman_plots}}
\end{figure}

@@

\noindent 
\begin{tabular}{@{}ll}
	cLBP & chronic low back pain\\
	CNN & convolutional neural network\\
	IMU & inertial measurement unit\\
	MS & multiple sclerosis\\
	OA & older adults\\
	PD & Parkinson's Disease\\
	TCN & temporal convolutional network\\
	YA & younger adults
\end{tabular}}

@@

%=====================================
% References, variant B: internal bibliography
%=====================================
\begin{thebibliography}{999}
	\bibitem[Snijders(2007)]{Snijders2007}
	Snijders,~A.H.; {van de Warrenburg},~B.P.; Giladi,~N.; Bloem,~B.R. Neurological gait disorders in elderly people: clinical approach and classification. {\em Lancet Neurol.} {\bf 2007}, {\em 6}(1), 63--74.
	\bibitem[Hodgins(2008)]{Hodgins2008}
	Hodgins,~D. The importance of measuring human gait. {\em Med. Device Technol.} {\bf 2008}, {\em 19}(5):42, 44-7.
	\bibitem[DelDin(2019)]{DelDin2019}
	{Del Din},~S.; Elshehabi,~M.; Galna,~B.; Hobert,~M.A.; Warmerdam,~E.; Suenkel,~U.; Brockmann,~K.; Metzger,~F.; Hansen,~C.; Berg,~D.; Rochester,~L.; Maetzler,~W. Gait analysis with wearables predicts conversion to Parkinson disease. {\em Ann. Neurol.} {\bf 2019}, {\em 86}(3), 357--367.
	\bibitem[Koenig(2017]{Koenig2017}
	K\"{o}nig,~A.; Klaming,~L.; Pijl,~M.; Demeurraux,~A.; Davis,~R.; Robert,~P. Objective measurement of gait parameters in healthy and cognitively impaired elderly using the dual-task paradigm. {\em Aging. Clin. Exp. Res.} {\bf 2017}, {\em 29}(6), 1181--1189.
	\bibitem[Bertoli(2018)]{Bertoli2018}
	Bertoli,~M.; Cereatti,~A.; Trojaniello,~D.; Avanzino,~L.; Pelosin,~E.; Del Din,~S.; Rochester,~L.; Ginis,~P.; Bekkers,~E.M.J.; Mirelman,~A.; Hausdorff,~J.M.; {Della Croce},~U. Estimation of spatio-temporal parameters of gait from magneto-inertial measurement units: multicenter validation among Parkinson, mildly cognitively impaired and healthy older adults. {\em Biomed. Eng. Online} {\bf 2018}, {\em 17}(58).
	\bibitem[SchroederVon(1995)]{SchroederVon1995}
	von Schroeder,~H.P.; Coutts,~R.D.; Lyden,~P.D.; Billings,~E.,~Jr; Nickel,~V. L. Gait parameters following stroke: a practical assessment. {\em J. Rehabil. Res. Dev.} {\bf 1995}, {\em 32}(1), 25--31.
	\bibitem[Mohan(2021)]{Mohan2021}
	Mohan,~D.M.; Khandoker,~A.H.; Wasti,~S.A.; Ismail~Ibrahim~Ismail~Alali,~S.; Jelinek,~H.F.; Khalaf,~K. Assessment methods of post-stroke gait: A scoping review of technology-driven approaches to gait characterization and analysis. {\em Front. Neurol.} {\bf 2021}, {\em 12}.
	\bibitem[Griskevicius(2016)]{Griskevicius2016}
	Gri\v{s}kevi\v{c}ius,~J.; Apanskien\.{e},~V.; \v{Z}i\v{z}ien\.{e},~J.; Daunoravi\v{c}ien\.{e},~K.; Ov\v{c}inikova,~A.; Kizlaitien\.{e},~R.; Sereik\.{e},~I.; Kaubrys,~G.; Pauk,~J.; Id\'{z}kowski,~A. Estimation of temporal gait parameters of multiple sclerosis patients in clinical setting using inertial sensors. In Proceedings of the 11th Int Conf BIOMDLORE 2016, Druskininkai, Lithuania, 20--22 Oct 2016; 80--82.
	\bibitem[Flachenecker(2019)]{Flachenecker2019}
	Flachenecker,~F.; Ga{\ss}ner,~H.; Hannik,~J.; Lee,~D.H.; Flachenecker,~P.; Winkler,~J.; Eskofier,~B.; Linker,~R.A.; Klucken,~J. Objective sensor-based gait measures reflect motor impairment in multiple sclerosis patients: Reliability and clinical validation of a wearable sensor device. {\em Mult. Scler. Relat. Dis.} {\bf 2019}, {\em 39}, 101903.
	\bibitem[Hannink(2016)]{Hannink2016}

>>>>

%=================================================================
\documentclass[journal,article,submit,pdftex,moreauthors]{Definitions/mdpi} 
% For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept". The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}. This is especially recommended for submission to arXiv, where line numbers should be removed before posting. For preprints.org, the editorial staff will make this change immediately prior to posting.

@@

%=================================================================
% Full title of the paper (Capitalized)
\Title{Title}

% MDPI internal command: Title for citation in the left column
\TitleCitation{Title}

% Author Orchid ID: enter ID or remove command
\newcommand{\orcidauthorA}{0000-0000-0000-000X} % Add \orcidA{} behind the author's name
%\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name

% Authors, for the paper (add full first names)
\Author{Firstname Lastname $^{1,\dagger,\ddagger}$\orcidA{}, Firstname Lastname $^{2,\ddagger}$ and Firstname Lastname $^{2,}$*}

@@

% MDPI internal command: Authors, for metadata in PDF
\AuthorNames{Firstname Lastname, Firstname Lastname and Firstname Lastname}

% MDPI internal command: Authors, for citation in the left column
\AuthorCitation{Lastname, F.; Lastname, F.; Lastname, F.}
% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Affiliation 1; e-mail@e-mail.com\\
$^{2}$ \quad Affiliation 2; e-mail@e-mail.com}

% Contact information of the corresponding author
\corres{Correspondence: e-mail@e-mail.com; Tel.: (optional; include country code; if there are multiple corresponding authors, add author initials) +xx-xxxx-xxx-xxxx (F.L.)}

% Current address and/or shared authorship
\firstnote{Current address: Affiliation 3.} 
\secondnote{These authors contributed equally to this work.}
% The commands \thirdnote{} till \eighthnote{} are available for further notes

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setcounter{section}{-1} %% Remove this when starting to work on the template.
\section{How to Use this Template}

The template details the sections that can be used in a manuscript. Note that the order and names of article sections may differ from the requirements of the journal (e.g., the positioning of the Materials and Methods section). Please check the instructions on the authors' page of the journal to verify the correct order and names. For any questions, please contact the editorial office of the journal or support@mdpi.com. For LaTeX-related questions please contact latex@mdpi.com.%\endnote{This is an endnote.} % To use endnotes, please un-comment \printendnotes below (before References). Only journal Laws uses \footnote.

% The order of the section titles is: Introduction, Materials and Methods, Results, Discussion, Conclusions for these journals: aerospace,algorithms,antibodies,antioxidants,atmosphere,axioms,biomedicines,carbon,crystals,designs,diagnostics,environments,fermentation,fluids,forests,fractalfract,informatics,information,inventions,jfmk,jrfm,lubricants,neonatalscreening,neuroglia,particles,pharmaceutics,polymers,processes,technologies,viruses,vision

\section{Introduction}

The introduction should briefly place the study in a broad context and highlight why it is important. It should define the purpose of the work and its significance. The current state of the research field should be reviewed carefully and key publications cited. Please highlight controversial and diverging hypotheses when necessary. Finally, briefly mention the main aim of the work and highlight the principal conclusions. As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research. Citing a journal paper \cite{ref-journal}. Now citing a book reference \cite{ref-book1,ref-book2} or other reference types \cite{ref-unpublish,ref-communication,ref-proceeding}. Please use the command \citep{ref-thesis,ref-url} for the following MDPI journals, which use author--date citation: Administrative Sciences, Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, Journal of Intelligence, Journalism and Media, JRFM, Languages, Laws, Religions, Risks, Social Sciences, Literature.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Materials and Methods}

Materials and Methods should be described with sufficient details to allow others to replicate and build on published results. Please note that publication of your manuscript implicates that you must make all materials, data, computer code, and protocols associated with the publication available to readers. Please disclose at the submission stage any restrictions on the availability of materials or information. New methods and protocols should be described in detail while well-established methods can be briefly described and appropriately cited.

Research manuscripts reporting large datasets that are deposited in a publicly avail-able database should specify where the data have been deposited and provide the relevant accession numbers. If the accession numbers have not yet been obtained at the time of submission, please state that they will be provided during review. They must be provided prior to publication.

Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.
\begin{quote}
This is an example of a quote.
\end{quote}

@@

@@

\noindent 
\begin{tabular}{@{}ll}
MDPI & Multidisciplinary Digital Publishing Institute\\
DOAJ & Directory of open access journals\\
TLA & Three letter acronym\\
LD & Linear dichroism
\end{tabular}}

@@

%=====================================
% References, variant B: internal bibliography
%=====================================
\begin{thebibliography}{999}
% Reference 1
\bibitem[Author1(year)]{ref-journal}
Author~1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
% Reference 2
\bibitem[Author2(year)]{ref-book1}
Author~2, L. The title of the cited contribution. In {\em The Book Title}; Editor 1, F., Editor 2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
% Reference 3
\bibitem[Author3(year)]{ref-book2}
Author 1, A.; Author 2, B. \textit{Book Title}, 3rd ed.; Publisher: Publisher Location, Country, 2008; pp. 154--196.
% Reference 4
\bibitem[Author4(year)]{ref-unpublish}
Author 1, A.B.; Author 2, C. Title of Unpublished Work. \textit{Abbreviated Journal Name} year, \textit{phrase indicating stage of publication (submitted; accepted; in press)}.
% Reference 5
\bibitem[Author5(year)]{ref-communication}
Author 1, A.B. (University, City, State, Country); Author 2, C. (Institute, City, State, Country). Personal communication, 2012.
% Reference 6
\bibitem[Author6(year)]{ref-proceeding}
Author 1, A.B.; Author 2, C.D.; Author 3, E.F. Title of presentation. In Proceedings of the Name of the Conference, Location of Conference, Country, Date of Conference (Day Month Year); Abstract Number (optional), Pagination (optional).
% Reference 7
\bibitem[Author7(year)]{ref-thesis}
Author 1, A.B. Title of Thesis. Level of Thesis, Degree-Granting University, Location of University, Date of Completion.
% Reference 8
\bibitem[Author8(year)]{ref-url}
Title of Site. Available online: URL (accessed on Day Month Year).
\end{thebibliography}
